---
title: "P8105 HW2"
author: "Abhishek Ajay (aa4266)"
date: "October 1, 2018"
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```
 
#Problem 1

##Data import and manipulation

```{r p1_data_import}
nyc_transit_data = 
  read_csv("./datasets/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```
##Further Questions

###Question 1

How many distinct stations are there?

```{r p1_q1}
#count(distinct(nyc_transit_data, line, station_name))
distinct(nyc_transit_data, line, station_name) %>%
  count() %>%
  knitr::kable()
```

###Question 2

How many stations are ADA compliant?

```{r p1_q2}
nyc_transit_data %>%
  filter(ada == TRUE) %>%
  distinct(line, station_name) %>%
  count() %>%
  knitr::kable()
```
###Question 3

What proportion of station entrances / exits without vending allow entrance?

```{r p1_q3}
n_stations_without_vending =
  nyc_transit_data %>%
  filter(vending == "NO", entry == TRUE) %>%
  count()/nrow(nyc_transit_data[which(nyc_transit_data$vending == "NO"),]) %>%
  knitr::kable()
```

##Reformating of the data

Here we reformat the data such the route number and the route name are distinct variables.

```{r p1_reformatting}
nyc_transit_data_reformatting =
  nyc_transit_data %>%
  gather(key = route_number, value = route_name, route1:route11) %>%
  separate(route_number, into = c("route_str", "route_number"), sep = 5) %>%
  select(-route_str)
  #filter(distinct(route_number, route_name))
```

##Further questions on the reformatted table

###Question 1 

How many distinct stations serve the A train?

```{r p1_reformatting_q1}
nyc_transit_data_reformatting %>%
  filter(route_name == "A") %>%
  distinct(line, station_name) %>%
  count() %>%
  knitr::kable()
```

###Question 2

Of the stations that serve the A train, how many are ADA compliant?

```{r p1_reformatting_q2}
nyc_transit_data_reformatting %>%
  filter(route_name == "A", ada == TRUE) %>%
  distinct(line, station_name) %>%
  count() %>%
  knitr::kable()
```

#Problem 2

Import and manipulation of the Mr. Trash Wheel Dataset.

```{r p2_data_import}
mr_trash_wheel_data =
  read_excel("./datasets/HealthyHarborWaterWheelTotals2017-9-26.xlsx", range = "A2:O258") %>%
  janitor::clean_names() %>%
  filter(is.na(x_1)) %>%
  select(-x_1) %>%
  filter(!is.na(dumpster)) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))
```

Import and manipulation of the precipitation data from 2016 and 2017

```{r p2_data_import_precp}
mr_trash_wheel_precip_data_2017 =
  read_excel("./datasets/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3, range = "A2:B15") %>% 
  janitor::clean_names() %>%
  filter(!is.na(total)) %>%
  select(month, total_precipitation = total)

mr_trash_wheel_precip_data_2017$year = 2017

mr_trash_wheel_precip_data_2016 =
  read_excel("./datasets/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4, range = "A2:B15") %>% 
  janitor::clean_names() %>%
  filter(!is.na(total)) %>%
  select(month, total_precipitation = total)

mr_trash_wheel_precip_data_2016$year = 2016

```
